---
editor_options:
  chunk_output_type: console
---

# NOTES

-   This file is *not* included in the [source package](https://r-pkgs.org/structure.html#sec-source-package) because of the [`.nopurl` suffix in its
    filename](https://pkgpurl.rpkg.dev/reference/purl_rmd.html#-rmd-files-excluded-from-purling).

-   The chunks below have to be manually executed in order to regenerate the package data.

-   Although the datasets below are saved as ["internal data" in `R/sysdata.rda`](https://r-pkgs.org/data.html#sec-data-sysdata), they can still be exported and
    documented (by documenting the dataset's quoted name in the main `.Rmd` source file – which only works when the dataset is also `@export`ed), something [not
    explicitly mentioned](https://coolbutuseless.github.io/2018/12/10/r-packages-internal-and-external-data/) in the book [R
    Packages](https://r-pkgs.org/data.html#sec-data-data). To do so, you first need to manually add the `export()` directive in the NAMESPACE file since
    roxygen2 won't add it automatically.

# Define data

## `pkg_config`

```{r}
pkg_config <-
  tibble::tibble(key = character(),
                 default_value = list(),
                 default_value_dynamic = character(),
                 require = logical(),
                 description = character()) |>
  tibble::add_row(key = "base_url",
                  default_value = list("https://gitlab.com/api/v4"),
                  require = TRUE,
                  description = paste0("Base URL of the GitLab v4 RESTful API server to be accessed when `base_url` is not explicitly provided to the ",
                                       "functions of this package.")) |>
  tibble::add_row(key = "id_project",
                  require = TRUE,
                  description = "GitLab project identifier to use when `id_project` is not explicitly provided to the functions of this package.") |>
  tibble::add_row(key = "token",
                  default_value_dynamic = 'Sys.getenv("GITLAB_COM_TOKEN")',
                  require = FALSE,
                  description = paste0("[GitLab access token](https://docs.gitlab.com/ee/api/rest/authentication.html) used for authentication when `token` is",
                                       "not explicitly provided to the functions of this package. A personal access token for gitlab.com can be created under",
                                       "[this link](https://gitlab.com/-/user_settings/personal_access_tokens)."))
```

# Write data

Save all the small data objects as a single internal file `R/sysdata.rda`. Note that when documenting them, they must be explicitly `@export`ed to be available
to package users.

```{r}
usethis::use_data(pkg_config,
                  internal = TRUE,
                  overwrite = TRUE,
                  compress = "xz",
                  version = 3L)
```
