---
output: pal::gitlab_document
---

```{r}
#| label: init
#| include: false

knitr::opts_knit$set(root.dir = getwd())

library(rlang,
        include.only = "%|%")
library(magrittr,
        include.only = c("%>%", "%<>%", "%T>%", "%!>%", "%$%"))
```

# `r pal::desc_get_field_safe("Package")`

```{r}
#| label: pkg-desc
#| child: !expr pkgsnip::snip_path("pkg-desc.Rmd")
```

In contrast to the [**gitlabr**](https://thinkr-open.github.io/gitlabr/) package, this package is built upon [httr2](https://httr2.r-lib.org/) and features a
thoroughly designed [`api_req()`](https://gitlab.rpkg.dev/reference/api_req) core function that easily allows to implement the RESTful API like it's meant to be
used, including light-weight ("cached") requests etc. The package is designed in a [purely
functional](https://en.wikipedia.org/wiki/Purely_functional_programming) way[^1]. However, note that gitlabr still covers many more API endpoints than this
package.

```{r}
#| label: pkg-doc
#| eval: !expr '!isTRUE(getOption("pal.build_readme.is_pkgdown"))'
#| results: asis
#| echo: false

pkgsnip::md_snip(id = "pkgdown_site") %>%
  paste0("## Documentation\n\n",
         "[![Netlify Status](https://api.netlify.com/api/v1/badges/3a09179f-b2d1-4860-bc1d-318a9fb48d1c/deploy-status)]",
         "(https://app.netlify.com/sites/gitlab-rpkg-dev/deploys)\n\n",
         .) |>
  pal::cat_lines()
```

[^1]: Thanks to [funky's canonicalized way to package configuration](https://funky.rpkg.dev/), setting necessary state globally like the API's base URL and
    access token is still easy for users and transparent for developers.

## Installation

```{r}
#| label: pkg-instl-dev
#| child: !expr pkgsnip::snip_path("pkg-instl-dev-gitlab.Rmd")
```

```{r}
#| label: pkg-usage
#| eval: !expr isTRUE(getOption("pal.build_readme.is_pkgdown"))
#| results: asis
#| echo: false

pkgsnip::md_snip(id = "pkg_usage") %>%
  paste0("## Usage\n\n", .) |>
  pal::cat_lines()
```

## Package configuration

```{r}
#| label: pkg-config
#| child: !expr pkgsnip::snip_path("funky-config.Rmd")
```

## Development

### R Markdown format

```{r}
#| label: pkgpurl
#| child: !expr pkgsnip::snip_path("pkgpurl.Rmd")
```

### Coding style

```{r}
#| label: pkg-code-style
#| child: !expr pkgsnip::snip_path("pkg-code-style.Rmd")
```
